---
title: "Minimum viable analysis"
author: "Dahn-Young Dong"
date: "10/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Mimimum Viable Analysis of - Population tree construction and spatially-explicit admixture mapping

Generally not from a wheel, to bottom, to whole car... but from a skateboard, scooter, to bike, to car... each is a self-contained with instant gratification...
really big input data: 1. long time analysis 2. challenging to work with. 3. minimal possible dataset to approach the question.

Use Pilot dataset of 17 individuals to...

```{r}
require(gstudio)
require(vcfR)
require(pegas)
```

### GSTUDIO- genetic proportions onto map

```{r}
#import
pilot.loci <- pegas::read.vcf("./input/rasy_ymf_main.vcf", from = 10000, to = 10999)
#reuiqre output
write.csv(pilot.loci, "./output/pilot.loci")
#use gstudio::read pop
pilot.pop <- gstudio::read_population("./output/pilot.loci", header = TRUE, type = "separated", locus.columns = 1:1000, delim = "/") #trouble
```
```

### PhyloNet- to build a consensus introgression tree, if the populations evolve tree-like at all
